Rodent trapping studies to understand the prevalence of zoonotic disease in West Africa: A scoping review.

### 2021-08-30 reports/manuscript_v2.Rmd

  * We summarised the number of studies, the year in which trapping occurred and the country in which they were conducted. In studies not reporting the year of trapping we used the year of publication as a proxy.
  * We used the GPS coordinates of single trap, trapping grid/line or study site and the number of trap nights to calculate trapping effort (trap night density) within level 2 administrative areas in West Africa. 
  * For studies not reporting the number of trap nights we imputed the number based on the median trap success rate from studies which reported the number of trap nights (matched to building or non-building based study sites). The median trap success rate for all rodents at a defined trap site was calculated for both trap sites including built environments and non-built environments. The number of rodents trapped at a trap site was then used in combination with these values of trap success to impute the number of trap nights for trap sites with no reported trap nights.
  * We summarised the habitat types of trap sites based on information reported in the study.
  * For the subset of studies investigating rodent zoonoses we compared the location of trapping sites was to SEDAC Global Population Density estimates (1970-2000 (10 year) and 2000-2020 (5 year)) and to land cover classifications (1975, 2000 and 2013) to explore the association of human population and rodent trapping effort.
  * We mapped the presence and absence of rodent species and compared this to the presence and absence reported by both GBIF and IUCN to give a measure of the extent of each species range in which they have been sampled. 
  * **Which statistical test to compare?** Ranked correlation to compare presence of habitat type against number of trap nights in habitat? Same analysis but weighted for human population?

@@

  * Rodent trapping took place across 1,331 trap sites with at least one trap site recorded from 14 West African countries. No rodent trapping studies were identified from Gambia or Togo.
  * Thirty-one (25%) studies reported trapping at a single study site, 46 (37%) studies trapped at between two and five study sites, the remaining 49 studies trapped at between six and 93 study sites. Trap sites were situated in 273 (19.3%) of 1414 level 2 administrative regions from the 14 West African nations. The areas with highest trapping activity included the capital cities of Niger (Niamey), Sierra Leone (Freetown), Senegal (Dakar), Mali (Bamako) and Ghana (Accra) and the largest cities of Ivory Coast (Abidjan) and Benin (Cotonou). Outside of these cities, Northern Senegal (Fatique, Thies, Saint-Louis and Kedougou), Southern Guinea (Kindia and Nzerejore), Edo and Osun States in Nigeria and Eastern Sierra Leone were the most intensively studied regions (see Figure 1 - Panel B).
  * No studies reported trap habitats with reference to a standardised habitat classification scheme. 
  * Extracted habitat types were grouped into 30 categories (see Supplementary Material 2 for the habitat dictionary). At least one habitat was recorded for 17,122 trap sites (95%), with two or more habitats for a single trap site recorded for 4,403 (24%) sites. Single trap sites could span multiple habitat types resulting in 22,202 distinct habitat and trap sites. 
  * The most commonly trapped sites were in or around buildings (29%), in areas described as the rodents “natural habitat” (24%), in agricultural areas (13%) (e.g. rice fields, palm plantations), forests (8%) and in the area surrounding buildings (4%).
  * Explore whether trap sites are more likely to have been conducted within high population regions or whether there is an association with trap location and habitat as obtained from remote sensing data (Need to do this. The remote sensing data will be matched to the trap dates there will be several years between the trapping and data this is compared to it seems impractical to try and get annual population estimates or land classification so this will be chosen pragmatically)

>>>>

  * We summarised the number of studies, the year in which trapping occurred and the country in which they were conducted. In studies not reporting the year of trapping we used the year of publication as a proxy.
  * We used the GPS coordinates of single trap, trapping grid/line or study site and the number of trap nights to calculate trapping effort (trap night density) within level 2 administrative areas in West Africa. 
  * For studies not reporting the number of trap nights we imputed the number based on the median trap success rate from studies which reported the number of trap nights (matched to building or non-building based study sites). The median trap success rate for all rodents at a defined trap site was calculated for both trap sites including built environments and non-built environments. The number of rodents trapped at a trap site was then used in combination with these values of trap success to impute the number of trap nights for trap sites with no reported trap nights.
  * We summarised the habitat types of trap sites based on information reported in the study.
  * For the subset of studies investigating rodent zoonoses we compared the location of trapping sites with 2005 SEDAC Global Population Density estimates, the median year studies were commenced, using a Generalised Additive Model (GAM) incorporating a spatial interaction term to investigate the association of number of trap nights and human population density. The model structure was specified as below:
    * *trap night density* ~ *Tweedie*(*log*(Population density (2005)) + (Longitude * Latitude))   
  * We repeated this approach for land cover classifications obtained from the European Space Agency Copernicus dataset (2005) using a GAM incorporating a spatial interaction term to investigate the association of trap nights and land cover.
  * We mapped the presence and absence of rodent species and compared this to the presence and absence reported by both GBIF and IUCN to give a measure of the extent of each species range in which they have been sampled.

@@

  * Rodent trapping took place across 1,331 trap sites with at least one trap site recorded from 14 West African countries. No rodent trapping studies were identified from Gambia or Togo.
  * Thirty-one (25%) studies reported trapping at a single study site, 46 (37%) studies trapped at between two and five study sites, the remaining 49 studies trapped at between six and 93 study sites. Trap sites were situated in 273 (19.3%) of 1414 level 2 administrative regions from the 14 West African nations. The areas with highest trapping activity included the capital cities of Niger (Niamey), Sierra Leone (Freetown), Senegal (Dakar), Mali (Bamako) and Ghana (Accra) and the largest cities of Ivory Coast (Abidjan) and Benin (Cotonou). Outside of these cities, Northern Senegal (Fatique, Thies, Saint-Louis and Kedougou), Southern Guinea (Kindia and Nzerejore), Edo and Osun States in Nigeria and Eastern Sierra Leone were the most intensively studied regions (see Figure 1 - Panel B).
  * Greater numbers of trap nights were conducted in areas with higher population densities (*R~2~* = 0.167). In particular Mauritania, Northern Senegal, Eastern Sierra Leone, Eastern Guinea and South West Nigeria were trapped at increased rates than would be expected based on their population densities. South West Nigeria, Northern Nigeria, Liberia, the Ivory Coast, Ghana, Niger and Burkina Faso were trapped at lower rates.
  * No studies reported trap habitats with reference to a standardised habitat classification scheme. 
  * Extracted habitat types were grouped into 30 categories (see Supplementary Material 2 for the habitat dictionary). At least one habitat was recorded for 17,122 trap sites (95%), with two or more habitats for a single trap site recorded for 4,403 (24%) sites. Single trap sites could span multiple habitat types resulting in 22,202 distinct habitat and trap sites. 
  * The most commonly trapped sites were in or around buildings (29%), in areas described as the rodents “natural habitat” (24%), in agricultural areas (13%) (e.g. rice fields, palm plantations), forests (8%) and in the area surrounding buildings (4%).
  * Explore whether trap sites are more likely to have been conducted within high population regions or whether there is an association with trap location and habitat as obtained from remote sensing data (Need to do this. The remote sensing data will be matched to the trap dates there will be several years between the trapping and data this is compared to it seems impractical to try and get annual population estimates or land classification so this will be chosen pragmatically)

```{r, figure_1}

cowplot::plot_grid(plotlist = )

```
